
---
title: "Exploratory Data Analysis of Chicago Crime Data"
output: html_document
---

## Introduction

This document presents an exploratory data analysis (EDA) of a subset of Chicago's crime dataset.

## Setup

```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)
library(ggplot2)
library(sf)
```

## Data Loading

```{r load-data}
# Convert 'Date' to datetime for easier manipulation
crimes$Date <- as.Date(crimes$Date, format="%m/%d/%Y %H:%M:%S")
```


## 1. Yearly Trends with Plot

```{r yearly-trends-plot}
# Yearly Trends
yearly_trends <- crimes %>% 
  mutate(Year = year(Date)) %>% 
  count(Year) %>% 
  arrange(Year)

# Plotting
ggplot(yearly_trends, aes(x = Year, y = n)) +
  geom_line() +
  geom_point() +
  labs(title = "Yearly Crime Trends in Chicago", x = "Year", y = "Number of Crimes")
```

## 2. Seasonal and Monthly Trends

```{r monthly-trends}
# Monthly Trends
monthly_trends <- crimes %>% 
  mutate(Month = month(Date)) %>% 
  count(Month) %>% 
  arrange(Month)

monthly_trends
```

## 3. Day of the Week Analysis

```{r day-of-week-analysis}
# Day of the Week Analysis
day_of_week_trends <- crimes %>% 
  mutate(DayOfWeek = wday(Date, label = TRUE)) %>% 
  count(DayOfWeek) %>% 
  arrange(DayOfWeek)

day_of_week_trends
```

## 4. Spatial Analysis

```{r spatial-analysis}
# Assuming Latitude and Longitude are available in the dataset
crimes_sf <- st_as_sf(crimes, coords = c("Longitude", "Latitude"), crs = 4326)

# Plot
ggplot(data = crimes_sf) +
  geom_sf() +
  coord_sf(xlim = c(-87.94011, -87.52414), ylim = c(41.64454, 42.02304)) +
  theme_minimal()
```

## 5. Crime Type Analysis

```{r crime-type-analysis}
# Crime Type Analysis
crime_type_analysis <- crimes %>% 
  group_by(Primary.Type) %>% 
  summarise(Count = n()) %>% 
  arrange(desc(Count))

crime_type_analysis
```

## Conclusion

This document provided a detailed exploratory data analysis of the Chicago crimes dataset, including yearly trends, seasonal variations, day of the week patterns, spatial distribution, and crime type frequencies.
