
---
title: "Exploratory Data Analysis of Chicago Crime Data"
output: html_document
---

## Introduction

This document presents an exploratory data analysis (EDA) of a subset of Chicago's crime dataset, along with data from Chicago's community areas.

## Data Loading and Preprocessing

```{r load-data, message=FALSE}
library(tidyverse)
library(lubridate)

# Load the data
crimes <- read.csv("Crimes_-_2001_to_Present.csv")
comm_areas <- read.csv("CommAreas.csv")

# Convert date columns to Date type in crimes
crimes$Date <- mdy_hms(crimes$Date)
crimes$Updated.On <- mdy_hms(crimes$Updated.On)

# Handling missing values
crimes <- na.omit(crimes)
```

## Descriptive Analysis

```{r descriptive-analysis}
# Basic statistics
summary(crimes)

# Frequency of different types of crimes
crime_types <- crimes %>% 
  group_by(Primary.Type) %>%
  summarise(Count = n()) %>%
  arrange(desc(Count))

crime_types
```

## Geospatial Analysis

```{r geospatial-analysis, message=FALSE}
library(ggplot2)
library(sf)

# Basic mapping of crime locations
ggplot(crimes, aes(x = Longitude, y = Latitude)) + 
  geom_point() + 
  theme_minimal()
```

## Temporal Analysis

```{r temporal-analysis}
# Trend over years
crimes %>% 
  count(year = year(Date)) %>%
  ggplot(aes(x = year, y = n)) +
  geom_line() +
  labs(title = "Crime Trends Over the Years", x = "Year", y = "Number of Crimes")
```

## Community Area Analysis

```{r community-area-analysis}
# Merge datasets
combined_data <- merge(crimes, comm_areas, by.x = "Community.Area", by.y = "AREA_NUMBE")

# Analysis by community area
community_analysis <- combined_data %>% 
  group_by(COMMUNITY) %>%
  summarise(Total_Crimes = n()) %>%
  arrange(desc(Total_Crimes))

community_analysis
```

## Crime Type Analysis

```{r crime-type-analysis}
# Analysis of crime types
crime_type_analysis <- crimes %>% 
  group_by(Primary.Type, Description) %>%
  summarise(Count = n()) %>%
  arrange(desc(Count))

crime_type_analysis
```

## Conclusion

This document provided a comprehensive exploratory data analysis of the Chicago crimes dataset, encompassing various aspects like descriptive statistics, geospatial, temporal, and community area-based analyses.
